Practical Optional Types for Clojure

### 2015-02-28 related-work.tex
ocaml

\paragraph{Multimethods} 
Millstein and Chambers~\cite{MS02}
describe Dubious, a simple statically typed core language including multimethods that
dispatch on the type of its arguments. They tackle a key challenge for statically typing
multimethods: ``it is possible for two modules containing arbitrary multimethods to typecheck
successfully in isolation but generate type errors when linked together.''~\cite{MS02}

@@

\paragraph{Record Types} 
O'Caml-style extensible record types have been the subject of extensive research 
(eg. Wand~\cite{Wan89}, Cardelli and Mitchell~\cite{CM91}, Harper and Pierce~\cite{HP91})
most inference

@@

Typed Racket is a statically typed sister language of Racket. It
attempts to preserve existing Racket idioms and aims to type check
existing Racket code by simply adding top level type annotations~\cite{Tob10}.

Typed Racket fully expands all macro calls before type checking~\cite{Tob10},
avoiding the complex semantics of type checking macro definitions, an ongoing research area summarised
by~\citet{Her10}. Typed Clojure also expands macros before type checking.

\paragraph{Variable-arity Polymorphism}

~\citet{STF09} invented a type system supporting variable-arity polymorphism  % doesn't fit on a line
a version of which is included in the current implementation of Typed Racket.
Their main innovation centres around \emph{dotted type variables}, which represent a heterogeneous sequence
of types. Dotted type variables allow \emph{non-uniform} variable-arity function types,
which are used to check definitions and usages of functions with non-trivial rest parameters

\paragraph{Java Interoperability in Statically Typed Languages}

Scala~\cite{OCD+} has nullable references for compatibility with Java.
Programmers must manually check for
\java{null} like in Java to avoid null-pointer exceptions.

% def checkNullOrEmpty(v:Seq[Any]):Boolean = {
%   return v.getClass
% }
% 
% checkNullOrEmpty(null) ;=> NPE
%
% null.getClass

>>>>

\paragraph{Multimethods} 
\citet{MS02} describe Dubious, a simple statically typed core language including multimethods that
dispatch on the type of its arguments. 
In Dubious methods declare the types of arguments that they expect which corresponds to
exclusively using \clj{class} as the dispatch function in Typed Clojure.

@@

\paragraph{Record Types} 
OCaml-style extensible record types (eg. \citet{Wan89}, \citet{CM91}, \citet{HP91})
are defined polymorphically in terms of row variables.
HMaps in Typed Clojure are instead designed to be used with subtyping.

@@
